Online Laterality Battery Pre-Processing - short version

### 2020-05-23 battery-preprocessing_v5_DB.Rmd
some cosmetic changes; also ensured CF task was scored correctly, which had effect on outlier detection (previously all scored correct); also made subject codes simple numeric to make it easier to track errors. In doing the latter, may have messed up with a search and replace, so needs more checking, which is inprogress!

---
title: "Online Laterality Battery Pre-Processing - short version"
author: "Adam Parker + Dorothy Bishop"
date: "updated April 22nd 2020"
output:
  pdf_document: default
  html_document: default
---
NB. 2 changes Apr 22nd by DB: 
a) all data now in a Data folder, so when reading need to specify that; 
b) *demographics* data frame renamed to *allsum*.

This R Markdown file prcoesses the various components of the Online Laterality Battery which is implemented online via Gorilla.sc.

@@

@@

#NB for original data setwd("~/Dropbox/Online LDT/300 Ps")
# libraries 
library(readr)
library(yarrr)
library(dplyr)
library(tidyr)
library(car)
library(ggplot2)
library(knitr)
require(readr)  # for read_csv()
require(dplyr)  # for mutate()
require(tidyr)  # for unnest()
require(purrr)  # for map(), reduce()
library(GGally)
require(stringr) # string matching
library(ggpirate)
library(corrr) #easier correlations!
```

First, the demographics is processed.This gives us a participant ID, age, gender, footedness, and handedness.

```{r demograhics, warning=FALSE}
demo_dat_R <- read_csv("./data/demographics_R.csv", na = c("empty", "NA")) # read in data
demo_dat_L <- read_csv("./data/demographics_L.csv", na = c("empty", "NA")) # read in data
demo_dat <- rbind(demo_dat_R, demo_dat_L)
# relabel 
demo_dat$subject <- demo_dat$`Participant Public ID`
demo_dat$subject <- as.factor(demo_dat$subject)
# count and print subjects
nsub <-length(levels(demo_dat$subject))
print(paste0("subjects: ", nsub))

# create data frame
allsum <- data.frame(matrix(ncol = 7, nrow = nsub))
  colnames(allsum) <- c("subject", "age", "gender", "footedness", "handedness", "bilingual", "strongest.language")
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(demo_dat$subject)[i] # find subject ID
  myrows <- which(demo_dat$subject==subname) # select rows for this subject
  tmp <- data.frame(demo_dat[myrows,])

    myrow <- myrow+1

    allsum$subject[myrow] <- subname
    allsum$age[myrow] <- as.numeric(tmp[tmp$Question.Key == "Age",]$Response)
    allsum$gender[myrow] <- tmp[tmp$Question.Key == "Gender",]$Response
    allsum$footedness[myrow] <- tmp[tmp$Question.Key == "footedness",]$Response
    allsum$handedness[myrow] <- tmp[tmp$Question.Key == "categorical_hand",]$Response
    allsum$bilingual[myrow] <- tmp[tmp$Question.Key == "bilingual",]$Response
    allsum$strongest.language[myrow] <- tolower(tmp[tmp$Question.Key == "bilingual-text",]$Response)

@@

# Here we summarise the age broken down by handedness and gender. This will be important for any write up. 
#### NOTE. THIS IS NOT REPRESENTATIVE OF THE FINAL DATASET
# age
allsum %>%
    group_by(handedness, gender) %>%
    summarize(avg = mean(age), n = n(), sd = sd(age))

#create a numerical code for handedness - can be used in plots to distinguish handedness groups (eg if pch set to handpch, get circles and triangles)
allsum$handpch <-1
allsum$handpch[allsum$handedness=='Left'] <-2
```

@@

```{r sightednesss, warning=FALSE}
miles_dat_R <- read_csv("./data/Miles_R.csv", na = c("empty", "NA")) # read in data
miles_dat_L <- read_csv("./data/Miles_L.csv", na = c("empty", "NA")) # read in data
miles_dat <- rbind(miles_dat_R, miles_dat_L)
#porta_dat_R <- read_csv("./dat2/Porta_R.csv", na = c("empty", "NA")) # read in data
porta_dat_L <- read_csv("./data/Porta_L.csv", na = c("empty", "NA")) # read in data
porta_dat <- rbind(porta_dat_L)
# add test ID
miles_dat$test_ID <- "miles"
porta_dat$test_ID <- "porta"
# merge the two
sight_dat <- rbind(miles_dat, porta_dat)
# relabel 
sight_dat$subject <- sight_dat$`Participant Public ID`
sight_dat$subject <- as.factor(sight_dat$subject)

@@

# create data frame
sightedness <- data.frame(matrix(ncol = 3, nrow = nsub))
  colnames(sightedness) <- c("subject", "miles", "porta")
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(sight_dat$subject)[i] # find subject ID
  myrows <- which(sight_dat$subject==subname) # select rows for this subject
  tmp <- data.frame(sight_dat[myrows,])

    myrow <- myrow+1

@@

```{r EHI, warning=FALSE}
EHI_R <- read_csv("./data/EHI_R.csv", na = "empty") # read in data
EHI_L <- read_csv("./data/EHI_L.csv", na = "empty") # read in data
EHI <- rbind(EHI_R, EHI_L)

# select only responses
EHI <- EHI[EHI$`Question Key` == "response-2-quantised" | EHI$`Question Key` == "response-3-quantised" | 
           EHI$`Question Key` == "response-4-quantised" | EHI$`Question Key` == "response-5-quantised" |
           EHI$`Question Key` == "response-6-quantised" | EHI$`Question Key` == "response-7-quantised" | 
           EHI$`Question Key` == "response-8-quantised" | EHI$`Question Key` == "response-9-quantised" |
           EHI$`Question Key` == "response-10-quantised" | EHI$`Question Key` == "response-11-quantised" | 
           EHI$`Question Key` == "response-12-quantised" | EHI$`Question Key` == "response-13-quantised" |
           EHI$`Question Key` == "response-14-quantised" | EHI$`Question Key` == "response-15-quantised" | 
           EHI$`Question Key` == "response-16-quantised" | EHI$`Question Key` == "response-17-quantised" |
           EHI$`Question Key` == "response-18-quantised" | EHI$`Question Key` == "response-19-quantised" | 
           EHI$`Question Key` == "response-20-quantised",]

# recode variabels
EHI$subject <- EHI$`Participant Public ID`
EHI$subject <- as.factor(EHI$subject)

@@

# score responses according to the original scoring method in Oldifeld (1971)
EHI$right_hand <- 0
EHI$left_hand <- 0
for (r in 1:nrow(EHI)) {
  if(EHI$Response[r] == 1){
    EHI$right_hand[r]= 2 
  } else {
    if(EHI$Response[r] == 2){
      EHI$right_hand[r]= 1
    } else {
      if(EHI$Response[r] == 4){
        EHI$left_hand[r]= 1
      } else {
        if(EHI$Response[r] == 5){
          EHI$left_hand[r]= 2 
        } else {
          (EHI$right_hand[r]= 1) & (EHI$left_hand[r]= 1)
        }
      }
    }
  }
}

@@

myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  myrow <- myrow+1
  subname <- levels(EHI$subject)[i] # find subject ID
  myrows <- which(EHI$subject==subname) # select rows for this subject
  tmp <- data.frame(EHI[myrows,])

    EHI2$subject[myrow] <- subname # add row for subject
    EHI2$right_hand[myrow] <- sum(tmp$right_hand) # add row for hand
    EHI2$left_hand[myrow] <- sum(tmp$left_hand) # add row for number of buttons pressed
  }

# This is a summary of the EHI
for_index2 <- 
  EHI2 %>% 
  group_by(subject) %>%
  mutate(index_EHI= ((right_hand-left_hand)/(right_hand + left_hand))*100)

# now create a data frame only for index 
EHI_index <- select(for_index2, "subject", "index_EHI")
# merge with allsum
allsum <- merge(allsum, EHI_index, by= "subject")

@@

eplot<-ggplot(allsum, aes(handpch,index_EHI, colour=gender)) +
  geom_jitter(width=.25,height=2)+
  ggtitle("Handedness index on the Edinburgh Handedness Inventory")
eplot
#This is not perfect, but gives and idea to check EHI is valid etc.

@@

#NB one supposed right-hander has index_EHI around -50. 
#I will recode this one, as this seems an error.

@@

```{r lexTALE, warning=FALSE}
# read data
lexTALE_L <- read_csv("./data/LexTale_L.csv", na = "empty")
lexTALE_R <- read_csv("./data/lexTale_R.csv", na = "empty")
lexTALE <- rbind(lexTALE_R, lexTALE_L)
# filter
lexTALE <- lexTALE[lexTALE$Attempt == 1,]
lexTALE <- lexTALE[lexTALE$display == "Task",]
# rename variables
lexTALE$subject <- lexTALE$`Participant Public ID`
lexTALE$accuracy <- lexTALE$Correct
lexTALE$condition <- lexTALE$ANSWER
# code as factors
lexTALE$subject <- as.factor(lexTALE$subject)
lexTALE$condition <- as.factor(lexTALE$condition)
  levels(lexTALE$condition) <- c("non-word", "word")

# create data frame
lexTale_score <- data.frame(matrix(ncol = 4, nrow = nsub))
  colnames(lexTale_score) <- c("subject", "Word", "nonWord", "lexTALE")
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(lexTALE$subject)[i] # find subject ID
  myrows <- which(lexTALE$subject==subname) # select rows for this subject
  tmp <- data.frame(lexTALE[myrows,])

    myrow <- myrow+1

@@

# remove those who didn't pass the LexTALE
allsum <- allsum[allsum$lexTALE >= 80,]
```

@@

Before doing anything with the data, we now pull in all the data from all nodes and combine them from a sub folder in the directory. Participants, identified by random letter strings, are stacked on top of each other. We start by reading the data and cutting out unwanted information.

```{r readraw,include= FALSE, warning=FALSE}
# Set directory
for (d in 1:2){
  data_file <- "./data/day1_combined.csv"

  if (d ==2){
    data_file <-  "./data/day2_combined.csv"
  }
  #we will assign raw data to all_dat2 (day2), but then reassign to all_dat1 on first run through loop, so we have correct data for each day
all_dat2 <- read.csv(data_file, na.strings = "NA",stringsAsFactors=F)
# cleaning - do operations that apply for all tasks
all_dat2 %>% filter(Attempt >= 1) # only answers 
# write as factors
all_dat2$subject <- as.factor(all_dat2$Participant.Public.ID)
all_dat2$RT <- as.numeric(as.character(all_dat2$Reaction.Time))
all_dat2$Trial <-as.numeric(all_dat2$Trial.Number) #useful if we want to restrict analysis to subset of trials so retain this variable
all_dat2$Task.Name <- as.factor(all_dat2$Task.Name)
all_dat2$word<-substring(all_dat2$word,66,(nchar(all_dat2$word)-4)) #remove format from written word stimuli
if (d==1)
{all_dat1 <- all_dat2}
}

all_dat2$word_OG<-NA #create dummy column just to make it easier to align columns
collist <- colnames(all_dat1)
all_dat2 <- all_dat2[,collist]

#Check for overlap in subjects between the two sessions
which(levels(all_dat1$subject) %in% levels(all_dat2$subject))
```

@@

```{r RToutlierfunction}
remove.outliers <- function(origdat,lowcut,zcut){
# Outlier removal put in separate function, which can be applied to different task RTs.
# 
# origdat is the dataframe with data for outlier removal, mycolname is name of variable 
# for outlier removal (usually RT), lowcut is a cutoff value below which data excluded
# (typically v fast responses that are implausible, e.g. 201 ms for RT)
# and zcut is the constant in Hoaglin-Iglewicz method (usually 2.2)

@@

  for (i in 1:nsub) { # loop through subjects
  subname <- levels(origdat$subject)[i] # find subject ID
  myrows <- which(origdat$subject==subname) # select rows for this subject
  #NB myrows is NOT a consecutive series, because done in blocks
  firstrow <-min(myrows)
  tmp <- data.frame(origdat[myrows,])

@@

 # Identify 25th and 75th quartiles, and the difference between them
          lower_quartile <- quantile(RTcorr, probs=0.25, na.rm="TRUE")
          upper_quartile <- quantile(RTcorr, probs=0.75, na.rm="TRUE")
          quartile_diff <- upper_quartile - lower_quartile
        # Outliers are defined as being below or above 2.2 times the quartile difference
          lower_limit <- lower_quartile - zcut*quartile_diff
          upper_limit <- upper_quartile + zcut*quartile_diff
        # create outlier variable
w1 <- which(origdat$RT[myrows]>upper_limit) #rows with outlier, nb RELATIVE to myrows range
origdat$outlier[myrows[w1]]<-1
}
 #For RT we just remove slow outliers, so ignore lower_limit
origdat$RT[origdat$outlier>0]<-NA #RT variable now has outliers excluded as NA
return(origdat)
}
```

@@

```{r maketaskdataframe, warning=FALSE}
make.df <- function(origdat,varlist,latlist){
nsub<-length(unique(origdat$subject))
# create compact data frame 
temp_dat <- data.frame(matrix(ncol = 4, nrow = nsub*2))
colnames(temp_dat) <- varlist
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(origdat$subject)[i] # find subject ID
  myrows <- which(origdat$subject==subname) # select rows for this subject
  latcol <- which(colnames(origdat) == varlist[2])
  RTcol <- which(colnames(origdat) == varlist[4])
  tmp <- data.frame(origdat[myrows,])

@@

    temp_dat$subject[myrow] <- subname # add row for subject
    temp_dat$side[myrow] <- latlist[j] # add row for side
    temp_dat$accurate[myrow] <- sum(tmp1$Correct,na.rm=TRUE) # add row for N accurate response
    temp_dat$p.corr[myrow] <- 100*mean(tmp1$Correct,na.rm=TRUE) # add row for %accurate responses - NB for chimeric only error is wrong emotion - measure is just a measure of choice. For RDT, can make error of selecting wrong side, so % correct is meaningful

@@

```{r plotdata_ttest, warning=FALSE}
dopirate <- function(origdat,task,measure){
# measure for each side
# plot the number of correct responses in each side
pirateplot(formula = X1 ~ side,
           data = origdat,
           main = measure,
           theme = 0,
           pal = "southpark", # southpark color palette
           bean.f.o = .0, # Bean fill
           point.o = .3, # Points
           inf.f.o = .7, # Inference fill
           inf.b.o = .8, # Inference border
           avg.line.o = 1, # Average line
           bar.f.o = .5, # Bar
           point.pch = 21,
           point.cex = .7,
           inf.method= "ci")
# plot L & R to get slope
q <- ggplot(data=origdat, aes(x=side, y=X1, group=subject, color=subject)) +
    geom_line() +
    geom_point() +
    ggtitle(paste0(task,": ",measure," by side"))
q + theme_bw() + theme(legend.position = "none")
#  t test
print(paste0(task,measure))
myt1<-t.test(origdat$X1~ origdat$side, paired = TRUE, alternative = "two.sided")
print(myt1)
}
```

@@

makeLI <- function(allsums,origdat,task,mycolname,polarity,mycutoff){
# we do this for accuracy first
# reformat data frame
keeps <- c("subject", "side", "X1") #X1 holds either accuracy or RT
LIdf <- origdat[keeps]
LIdf <- spread(LIdf, side, X1)
# calculate each participants' lat index
LIdf <- 
  LIdf %>% 
  group_by(subject) %>%
  mutate(thisLI= ((Right - Left)/(Right + Left))*100)

LIdf$LI <- LIdf$thisLI*polarity #polarity can be set to reverse sign - eg for RT a large score is bad, whereas for acc it is good, so can have opp polarity

# check if accuracy or RT level is outside acceptable cutoff.
# Do this based on average for left and right
# LI will be set to NA for those outside limits
LIdf$avg <- (LIdf$Left+LIdf$Right)/2
w<-which(LIdf$avg<mycutoff)
if(length(w)>0){
LIdf$LI[w] <- NA}

# now create a data frame for lat index 
LIdf <- select(LIdf, "subject", "LI")
# now merge with allsum
allsums <- merge(allsums, LIdf, by= "subject")
lastcol<-length(colnames(allsums))
colnames(allsums)[lastcol]<-mycolname

@@

# now create a data frame for lat index 
Zdf <- select(Zdf, "subject", "Z")
# now merge with allsum
allsums <- merge(allsums, Zdf, by= "subject")
lastcol<-length(colnames(allsums))
colnames(allsums)[lastcol]<-mycolname

@@

This part of the R Markdown file prcoesses the "Chimeric Face Task' implmented online via Gorilla.sc which aims to assess the lateralisation of emotion recognition in a face processing task. In the task, participants are shown a chimeric face. Participants then respond by indicating the emotion that they felt was most strong expressed.

This task is near identical to that reported in Karlsson, Johnstone, and Carey (2019). Stimuli were kindly provided by Dr Michael Burt (https://www.dur.ac.uk/psychology/staff/?id=1942) and were symmetrical average images created from four male and four female faces (see Burt & Perrett, 1997; Innes et al., 2016).

```{r read_face_dat, warning=FALSE}
for (d in 1:2){ #we will process day 1 and day 2 together
  origfile <- all_dat1
  if (d==2)
  {origfile <- all_dat2}
#As with earlier reading in, we default to saving CF_dat2 (day2), but at end of the chunk we reassign to CF_day1 in the first run through the d loop
task <- "Chimeric faces"
wanted <- c("subject", "stimuli", "Response", "RT", "l_hemiface", "r_hemiface") # select wanted columns
disp <- "e_Task"
CF_dat2 <- procdata(origfile,wanted,task,disp) #use generic function (see above) to read in relevant columns

# count subjects
nsub <-length(levels(CF_dat2$subject))
print(paste('day',d,task,": subjects: ", nsub))

>>>>

---
title: "Online Laterality Battery Pre-Processing - short version"
author: "Adam Parker + Dorothy Bishop"
date: "updated May 21st 2020 (now notebook format)"
output: html_notebook
---
22 May 2020: AP made changes to N trials needed for inclusion  
NB. 2 changes Apr 22nd by DB: 
a) all data now in a Data folder, so when reading need to specify that; 
b) *demographics* data frame renamed to *allsum*.

This R Markdown file processes the various components of the Online Laterality Battery which is implemented online via Gorilla.sc.

@@

NB. Just also putting here relevant references that need to be looked at!
https://pubmed.ncbi.nlm.nih.gov/18050003/

@@

#NB for original data setwd("~/Dropbox/Online LDT/300 Ps")
# libraries 
library(readr)
library(yarrr)
library(dplyr)
library(tidyr)
library(car)
library(ggplot2)
library(knitr)
require(readr)  # for read_csv()
require(dplyr)  # for mutate()
require(purrr)  # for map(), reduce()
library(GGally)
require(stringr) # string matching
library(ggpirate)
library(corrr) #easier correlations!
```

First, the demographics is processed.This gives us a participant subject, age, gender, footedness, and handedness.

```{r demograhics, warning=FALSE}
#Reading from OSF - this only works if files are public

demo_dat <- read.csv('https://osf.io/7smj6/download',stringsAsFactor=F)

# relabel 
demo_dat$subject <- demo_dat$Participant.Public.subject
demo_dat$subject <- as.factor(demo_dat$subject)
# count and print subjects
nsub <-length(levels(demo_dat$subject))
print(paste0("subjects: ", nsub))

# create data frame
allsum <- data.frame(matrix(ncol = 7, nrow = nsub))
  colnames(allsum) <- c("subject", "age", "gender", "footedness", "handedness", "bilingual", "strongest.language")
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(demo_dat$subject)[i] # find subject subject
  myrows <- which(demo_dat$subject==subname) # select rows for this subject
  tmp <- data.frame(demo_dat[myrows,])

    myrow <- i

    allsum$subject[myrow] <- subname
    allsum$age[myrow] <- tmp[tmp$Question.Key == "Age",]$Response
    allsum$gender[myrow] <- tmp[tmp$Question.Key == "Gender",]$Response
    allsum$footedness[myrow] <- tmp[tmp$Question.Key == "footedness",]$Response
    allsum$handedness[myrow] <- tmp[tmp$Question.Key == "categorical_hand",]$Response
    allsum$bilingual[myrow] <- tmp[tmp$Question.Key == "bilingual",]$Response
    allsum$strongest.language[myrow] <- tolower(tmp[tmp$Question.Key == "bilingual-text",]$Response)

@@

#Check ages
range(allsum$age)
# Here we summarise the age broken down by handedness and gender. This will be important for any write up. 
#### NOTE. THIS IS NOT REPRESENTATIVE OF THE FINAL DATASET
# age
allsum %>%
    group_by(handedness, gender) %>%
    summarize(avg = mean(age), n = n(), sd = sd(age))

#create a numerical code for handedness - can be used in plots to distinguish handedness groups (eg if pch set to handpch, get circles and triangles)
allsum$handpch <-1  #numeric code for plotting handedness
allsum$handpch[allsum$handedness=='Left'] <-2
```

@@

```{r sightednesss, warning=FALSE}
porta_dat <- read.csv("https://osf.io/zmqkt/download",stringsAsFactors = F)
miles_dat <-  read.csv("https://osf.io/ekqsp//download",stringsAsFactors = F)

miles_dat$test_subject <- "miles"
porta_dat$test_subject <- "porta"
# merge the two
sight_dat <- rbind(miles_dat, porta_dat)
# relabel 
sight_dat$subject <- sight_dat$Participant.Public.subject
sight_dat$subject <- as.factor(sight_dat$subject)

@@

# create data frame
sightedness <- data.frame(matrix(ncol = 3, nrow = nsub))
  colnames(sightedness) <- c("subject", "miles", "porta")
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(sight_dat$subject)[i] # find subject subject
  myrows <- which(sight_dat$subject==subname) # select rows for this subject
  tmp <- data.frame(sight_dat[myrows,])

    myrow <- i

@@

```{r EHI, warning=FALSE}

EHI <- read.csv("https://osf.io/kfdma/download",stringsAsFactors = F)

# select only responses
EHI <- EHI[EHI$Question.Key == "response-2-quantised" | EHI$Question.Key == "response-3-quantised" | 
           EHI$Question.Key == "response-4-quantised" | EHI$Question.Key == "response-5-quantised" |
           EHI$Question.Key == "response-6-quantised" | EHI$Question.Key == "response-7-quantised" | 
           EHI$Question.Key == "response-8-quantised" | EHI$Question.Key == "response-9-quantised" |
           EHI$Question.Key == "response-10-quantised" | EHI$Question.Key == "response-11-quantised" | 
           EHI$Question.Key == "response-12-quantised" | EHI$Question.Key == "response-13-quantised" |
           EHI$Question.Key == "response-14-quantised" | EHI$Question.Key == "response-15-quantised" | 
           EHI$Question.Key == "response-16-quantised" | EHI$Question.Key == "response-17-quantised" |
           EHI$Question.Key == "response-18-quantised" | EHI$Question.Key == "response-19-quantised" | 
           EHI$Question.Key == "response-20-quantised",]

# recode variabels
EHI$subject <- EHI$Participant.Public.subject
EHI$subject <- as.factor(EHI$subject)

@@

# score responses according to the original scoring method in Oldfield (1971)
EHI$right_hand <- 0
EHI$left_hand <- 0
for (r in 1:nrow(EHI)) {
  if(EHI$Response[r] == 1){
    EHI$right_hand[r]= 2 
  } else {
    if(EHI$Response[r] == 2){
      EHI$right_hand[r]= 1
    } else {
      if(EHI$Response[r] == 4){
        EHI$left_hand[r]= 1
      } else {
        if(EHI$Response[r] == 5){
          EHI$left_hand[r]= 2 
        } else {
          (EHI$right_hand[r]= 1) & (EHI$left_hand[r]= 1)
        }
      }
    }
  }
}

@@

for (i in 1:nsub) { # loop through subjects
  myrow <- i
  subname <- levels(EHI$subject)[i] # find subject subject
  myrows <- which(EHI$subject==subname) # select rows for this subject
  tmp <- data.frame(EHI[myrows,])

    EHI2$subject[myrow] <- subname # add row for subject
    EHI2$right_hand[myrow] <- sum(tmp$right_hand) # sum for R hand
    EHI2$left_hand[myrow] <- sum(tmp$left_hand) # sum for L hand
  }

EHI2$index_EHI <- 100*(EHI2$right_hand-EHI2$left_hand)/(EHI2$right_hand+EHI2$left_hand)

# now create a data frame only for index 
# merge with allsum
allsum$index_EHI <- NA #initialise
for (i in 1:nrow(allsum)){
  thissub <- allsum$subject[i]
  w<-which(EHI2$subject==thissub)
 allsum$index_EHI[i] <- EHI2$index_EHI[w]
}

@@

eplot<-ggplot(allsum, aes(handpch,index_EHI, colour=gender)) +
  geom_jitter(wsubjectth=.25,height=2)+
  ggtitle("Handedness index on the Edinburgh Handedness Inventory")
eplot
#This is not perfect, but gives and subjectea to check EHI is valid etc.

@@

#NB two supposed right-handers have index_EHI around -50. 
#I will recode these, as this seems an error.

@@

```{r lexTALE, warning=FALSE}
# read data
lexTALE <- read.csv("https://osf.io/h4btp/download",stringsAsFactors = F)

# filter
lexTALE <- lexTALE[lexTALE$Attempt == 1,]
lexTALE <- lexTALE[lexTALE$display == "Task",]
# rename variables
lexTALE$subject <- lexTALE$Participant.Public.ID
lexTALE$accuracy <- lexTALE$Correct
lexTALE$condition <- lexTALE$ANSWER
# code as factors
lexTALE$subject <- as.factor(lexTALE$subject)
lexTALE$condition <- as.factor(lexTALE$condition)
  levels(lexTALE$condition) <- c("non-word", "word")

# create data frame
lexTale_score <- data.frame(matrix(ncol = 4, nrow = nsub))
  colnames(lexTale_score) <- c("subject", "Word", "nonWord", "lexTALE")

for (i in 1:nsub) { # loop through subjects
  subname <- levels(lexTALE$subject)[i] # find subject subject
  myrows <- which(lexTALE$subject==subname) # select rows for this subject
  tmp <- data.frame(lexTALE[myrows,])

    myrow <- i

@@

# remove those who dIDn't pass the LexTALE
allsum <- allsum[allsum$lexTALE >= 80,]
```

@@

Before doing anything with the data, we now pull in all the data from all nodes and combine them from a sub folder in the directory. Participants, IDentified by random letter strings, are stacked on top of each other. We start by reading the data and cutting out unwanted information.

```{r readraw,include= FALSE, warning=FALSE}
# Set directory
for (d in 1:2){
  data_file <- "https://osf.io/mhk29/download" #day 1 data

  if (d ==2){
    data_file <-  "https://osf.io/ebm6s/download"
  }
  #we will assign raw data to all_dat2 (day2), but then reassign to all_dat1 on first run through loop, so we have correct data for each day
all_dat2 <- read.csv(data_file, na.strings = "NA",stringsAsFactors=F)
# cleaning - do operations that apply for all tasks
all_dat2 %>% filter(Attempt >= 1) # only answers 
# write as factors
all_dat2$subject <- as.factor(all_dat2$Participant.Public.ID)
all_dat2$RT <- as.numeric(as.character(all_dat2$Reaction.Time))
all_dat2$Trial <-as.numeric(all_dat2$Trial.Number) #useful if we want to restrict analysis to subset of trials so retain this variable
all_dat2$Task.Name <- as.factor(all_dat2$Task.Name)
all_dat2$word<-substring(all_dat2$word,66,(nchar(all_dat2$word)-4)) #remove format from written word stimuli
if (d==1)
{all_dat1 <- all_dat2}
}

all_dat2$word_OG<-NA #create dummy column just to make it easier to align columns
collist <- colnames(all_dat1)
all_dat2 <- all_dat2[,collist]
```

Create a simple numerical ID to make life easier if we need to identify datapoints which are odd in some way. Ensure this is consistent across all_dat1 and all_dat2 and allsum.
We will rename the current subject variable so we can reuse this one  for all future computations.

```{r makesubject}
#Use allsum as the basis for the check
#First rename all subject to origID
allsum <- allsum %>% rename(origID = subject) 
all_dat1 <- all_dat1 %>% rename(origID = subject) 
all_dat2 <- all_dat2 %>% rename(origID = subject) 
for (n in 1:nrow(allsum)){
  allsum$subject <- n
  w<-which(all_dat1$origID == allsum$origID[n])
  all_dat1$subject <-n
  w<-which(all_dat2$origID == allsum$origID[n])
  all_dat2$subject <-n
}
```

@@

Outlier removal put in separate function, which can be applied to different task RTs.  
Origdat is the dataframe with data for outlier removal, mycolname is name of variable  
For outlier removal (usually RT), lowcut is a cutoff value below which data excluded (typically v fast responses that are implausible, e.g. 201 ms for RT)
 and zcut is the constant in Hoaglin-Iglewicz method (usually 2.2)

```{r RToutlierfunction}
detect.outliers <- function(origdat,lowcut,zcut){

@@

  for (i in 1:nsub) { # loop through subjects
  subname <- levels(origdat$subject)[i] # find subject 
  myrows <- which(origdat$subject==subname) # select rows for this subject
  #NB myrows is NOT a consecutive series, because done in blocks
  firstrow <-min(myrows)
  tmp <- data.frame(origdat[myrows,])

@@

 # IDentify 25th and 75th quartiles, and the difference between them
          lower_quartile <- quantile(RTcorr, probs=0.25, na.rm="TRUE")
          upper_quartile <- quantile(RTcorr, probs=0.75, na.rm="TRUE")
          quartile_diff <- upper_quartile - lower_quartile
        # Outliers are defined as being below or above 2.2 times the quartile difference
          lower_limit <- lower_quartile - zcut*quartile_diff
          upper_limit <- upper_quartile + zcut*quartile_diff
        # create outlier variable
w1 <- which(origdat$RT[myrows]>upper_limit) #rows with outlier, nb RELATIVE to myrows range
origdat$outlier[myrows[w1]]<-1
}
 #For RT we just remove slow outliers, so ignore lower_limit
origdat$RT[origdat$outlier>0]<-NA #RT variable now has outliers excluded as NA
return(origdat)
}
```

@@

```{r maketaskdataframe, warning=FALSE}
make.df <- function(origdat,varlist,latlist){
nsub<-length(unique(origdat$subject))
# create compact data frame 
temp_dat <- data.frame(matrix(ncol = 4, nrow = nsub*2))
colnames(temp_dat) <- varlist
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(origdat$subject)[i] # find subject subject
  myrows <- which(origdat$subject==subname) # select rows for this subject
  latcol <- which(colnames(origdat) == varlist[2])
  RTcol <- which(colnames(origdat) == varlist[4])
  tmp <- data.frame(origdat[myrows,])

@@

    temp_dat$subject[myrow] <- subname # add row for subject
    temp_dat$side[myrow] <- latlist[j] # add row for sIDe
    temp_dat$accurate[myrow] <- sum(tmp1$Correct,na.rm=TRUE) # add row for N accurate response
    temp_dat$p.corr[myrow] <- 100*mean(tmp1$Correct,na.rm=TRUE) # add row for %accurate responses - NB for chimeric only error is wrong emotion - measure is just a measure of choice. For RDT, can make error of selecting wrong sIDe, so % correct is meaningful

@@

```{r plotdata_ttest, warning=FALSE}
dopirate <- function(origdat,task,measure){
# measure for each sIDe
# plot the number of correct responses in each sIDe
  myhead<-paste0(task,": ",measure)
pirateplot(formula = X1 ~ side,
           data = origdat,
           main = myhead,
           theme = 0,
           pal = "southpark", # southpark color palette
           bean.f.o = .0, # Bean fill
           point.o = .3, # Points
           inf.f.o = .7, # Inference fill
           inf.b.o = .8, # Inference border
           avg.line.o = 1, # Average line
           bar.f.o = .5, # Bar
           point.pch = 21,
           point.cex = .7,
           inf.method= "ci")
# plot L & R to get slope
q <- ggplot(data=origdat, aes(x=side, y=X1, group=subject, color=subject)) +
    geom_line() +
    geom_point() +
    ggtitle(paste0(task,": ",measure," by side"))
q + theme_bw() + theme(legend.position = "none")
#  t test
print(paste0(task,measure))
myt1<-t.test(origdat$X1~ origdat$side, paired = TRUE, alternative = "two.sided")
print(myt1)
}
```

@@

makeLI <- function(allsum,origdat,task,mycolname,polarity,mycutoff){
# we do this for accuracy first
# reformat data frame
keeps <- c("subject", "side", "X1") #X1 holds either accuracy or RT
Lsubjectf <- origdat[keeps]
Lsubjectf <- spread(Lsubjectf, side, X1)
# calculate each participants' lat index

Lsubjectf$thisLI <- 100*(Lsubjectf$Right - Lsubjectf$Left)/(Lsubjectf$Right + Lsubjectf$Left)

Lsubjectf$LI <- Lsubjectf$thisLI*polarity #polarity can be set to reverse sign - eg for RT a large score is bad, whereas for acc it is good, so can have opp polarity

# check if accuracy or RT level is outside acceptable cutoff.
# Do this based on average for left and right
# LI will be set to NA for those outside limits
Lsubjectf$avg <- (Lsubjectf$Left+Lsubjectf$Right)/2
w<-which(Lsubjectf$avg<mycutoff)
if(length(w)>0){
Lsubjectf$LI[w] <- NA}

# now create a data frame for lat index 
Lsubjectf <- select(Lsubjectf, "subject", "LI")
# now merge with allsum
# First check if this column already exists and delete it if so
w<-which(colnames(allsum)==mycolname)
allsum<-allsum[,-w]
#now add this LI
allsum <- merge(allsum, Lsubjectf, by= "subject")
lastcol<-length(colnames(allsum))
colnames(allsum)[lastcol]<-mycolname

return(allsum)
}
```

@@

# now create a data frame for lat index 
Zdf <- select(Zdf, "subject", "Z")
# now merge with allsum
# First check if this column already exists and delete it if so
w<-which(colnames(allsums)==mycolname)
allsums<-allsums[,-w]

allsums <- merge(allsums, Zdf, by= "subject")
lastcol<-length(colnames(allsums))
colnames(allsums)[lastcol]<-mycolname

@@

This part of the notebook processes the "Chimeric Face Task' implmented online via Gorilla.sc which aims to assess the lateralisation of emotion recognition in a face processing task. In the task, participants are shown a chimeric face. Participants then respond by indicating the emotion that they felt was most strong expressed.

This task is near identical to that reported in Karlsson, Johnstone, and Carey (2019). Stimuli were kindly provided by Dr Michael Burt (https://www.dur.ac.uk/psychology/staff/?subject=1942) and were symmetrical average images created from four male and four female faces (see Burt & Perrett, 1997; Innes et al., 2016). Stimuli for each combination of anger, disgust, happy and sad (8 left-first, 8 right-first) plus 8 of each emotion where both sides the same the same to check overall perception. Total of 96 chimeric stimuli presented in each session. We will exclude those who are inaccurate when both faces the same (as in preregistration). Responses to chimeric faces can be coded as whether they correspond to emotion on L or R. Note that if the reported emotion does not match either side, then trial is not counted.

The chunk below also detects outliers, and codes them as 1 if the RT is outside the range specified in Hoaglin-Iglewicz, and as 2 if the RT is < 200. Outliertable shows each subject with Ns with different outlier codes. Some people have 0-1 trials - these seem to be ones where the task was skipped for some reason - I have checked and they seem to have no data.
One person has twice the number of trials in one session. We'll deal with that by just taking the first 96 trials.
